---
title: "Analysis Inhaler"
author: "Joana"
date: today
toc: true
abstract-title: "Research question:"
abstract: "Does a new inhaled corticosteroid (ICS-X) perform better in a Randomized control trial over 6 weeks in patients with mild persistent asthma compared to a standard inhaler?"
format:
  html:
    embed-resources: true
    number-sections: true
    code-fold: show
editor: visual
editor_options:
  chunk_output_type: console
execute:
  warning: false
  message: false
---

## Prepare session
In this step the Quarto project has been prepared for the analysis of the data set lung function which contains 180 data points of distinct patients. The data set displays information about the age, FEV-1 improvement, baseline FEV-1, treatment group and smoking status. 

```{r}
#| label: prepare-session
#| echo: false

# Load packages
  library(pastecs)
  library(ggpubr); theme_set(theme_bw())
  library(GGally)
  library(performance)
  library(MuMIn)
  library(ggeffects)

  library(knitr) 

setwd("/Users/joanawick/Desktop/BME-321-FS2026-B")

# Import data
lung <- read.csv("data/processed/lung-function.csv")

```


## Explore the data
To look at the different inhalation methods, the dataset is analyzed numerically and graphically.

### Numerically
For the numerical exploration of the data, first the tabel @tbl-explore-data has been created, with which the descriptive statistics can be looked at. 
```{r}
#| label: tbl-explore-data
#| tbl-cap: "Descriptive statistics table of the variables of interest."
#| echo: false

# Numerically
 sum_stats<-
   stat.desc(lung [, c("fev1_improvement", "age", "baseline_fev1", "treatment_group")], norm= T)

 kable(sum_stats, digits= 1)


```

To get information about the proportion in which the various treatments have been used, the tabel @tbl-treatment-group has been created.
```{r}
#| label: tbl-treatment-group
#| tbl-cap: "This table displays information about the proportion of the treatment groups."
#| echo: false

## numerically
table(lung$treatment_group)
```

Another important factor in the analysis is the smoking status. The @tbl-smoking-status serves to further elaborate the smoking status of the participants.
```{r}
#| label: tbl-smoking-status
#| tbl-cap: "This table displays information about the proportion of the participants which smoke."
#| echo: false

## numerically
table(lung$smoking_status)
```


### Graphically 
To get a further impression of the data, it can be displayed graphically. The first step is to compare the treatment groups with @fig-boxplot.

```{r}
#| label: fig-boxplot
#| fig-cap: "This figure shows the outcome between the treatment groups."
#| echo: false 

ggplot(lung, aes(x = treatment_group, y = fev1_improvement)) +
  geom_boxplot() +
  labs(title = "Improvement by Treatment Groups",
    x = "Treatment Group", y = "FEV1 Improvement")

```

With the second @fig-histogram, the distribution of the FEV1 improvement can be explored.
```{r}
#| label: fig-histogram
#| fig-cap: "This figure shows how the FEV1 improvement is distributed."
#| echo: false 

ggplot(lung, aes(x = fev1_improvement)) +
  geom_histogram(bins = 20) +
  labs(title = "Distribution FEV1 Improvement",
    x = "FEV1 Improvement", y = "Count")

```

## Fit models 
After the initial graphical and numerical exploration of the data, the models have been fit (@tbl-fit-models). As we have a continous outcome variable (FEV1 improvement), the data is fitted with a linear model. 

### Formulate and compare
```{r}
#| label: tbl-fit-models
#| tbl-cap: "Model comparison."
#| echo: true
#| results: "hide"

# fit models
m0 <- lm(fev1_improvement ~ 1, lung)
m1 <- lm(fev1_improvement ~ treatment_group, lung)
m2 <- lm(fev1_improvement ~ treatment_group + age + baseline_fev1, lung)
m3 <- lm(fev1_improvement ~ treatment_group + age + baseline_fev1 + smoking_status, lung)

# compare models
kable(compare_performance(m0, m1, m2, m3), digits= 3)

```

After fitting and comparing the models, m3 has been chosen to continue as it has the lowest AIC and BIC and the highest adjusted R^2. M3 has an adjusted R^2 has a `r round(performance::r2(m3)$R2_adjusted, 3)`.

### Validate the preferred model
Before the chosen model can be interpreted, it has to be validated. To do so, the @fig-model-diagnostic has been created. Model 3 works well to continue with.

```{r}
#| label: fig-model-diagnostic
#| fig-cap: "This figures show the diagonstic plots for the preferred linear model (m3)."

check_model(m3)
```

## Interpret model
As m3 has been chosen and validated, the data interpretation can continue with it. 

```{r}
#| label: interpret-obj
#| include: false

# summary
summary(m3)

# coefficient table
coef.table <- round(summary(m3)$coefficients, 3)

# adjusted R2
r2 <- round(summary(m3)$adj.r.squared, 3)

# model AIC
model_aic <- round(AIC(m3), 2)

# pred
pred_age <- predict_response(m3, "age")
pred_baseline <- predict_response(m3, "baseline_fev1")
pred_treatment <- predict_response(m3, "treatment_group")

```

```{r}
#| label: fig-pred-age
#| fig-cap: "Predicted FEV1 improvement at different ages."
#| echo: false 

ggplot(pred_age, aes(x, predicted)) +
  geom_point(data = lung, aes(age, fev1_improvement), alpha = 0.5) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) +
  geom_line() +
  labs(x = "Age", y = "predicted FEV1 improvement")

```

```{r}
#| label: fig-pred-baseline
#| fig-cap: "Predicted improvement with FEV1 across the baseline."
#| echo: false 

ggplot(pred_baseline, aes(x, predicted)) +
  geom_point(data = lung, aes(baseline_fev1, fev1_improvement), alpha = 0.5) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) +
  geom_line() +
  labs(x = "Baseline FEV1", y = "predicted FEV1 improvement")
```

```{r}
#| label: fig-pred-treatment
#| fig-cap: "Predicted improvement with FEV1 in the different treatment groups."
#| echo: false 

ggplot(pred_treatment, aes(x, predicted)) +
  geom_point(data = lung, aes(treatment_group, fev1_improvement), alpha = 0.5) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) +
  geom_line() +
  labs(x = "Treatment group", y = "predicted FEV1 improvement")
```

```{r}
#| label: tbl-model-results
#| tbl-cap: !expr paste("Linear model results for the preferred model. Adjusted R² = ", r2, ", AIC = ", model_aic, ".")
#| echo: false 

kable(coef.table)
```

## Conclusion
The linear model (m3) which has been chosen explains `r round(r2 * 100, 1)`% of the variability in the improvement with the FEV1 treatment. Overall the patients with the standard treatment showed significantly lower (-0.063, p = 0.016) improvement. This leads to the conclusion the ICSX helps improving the lung function.
Looking at the FEV1 improvenment at different ages, showed that it's reduced with higher age (-0.009, p < 0.001). Higher baseline of FEV1 led also to a less good improvement (-0.182, p < 0.001). This could be because there is less room for improvement. 
Looking if the participants smoke or not also lead to an interesting conclusion. Former smokers and participants who never smoked had a lower FEV1 improvement compared to smokers. The strongest reduction reduction have never smokers (-0.213, p < 0.001). 


```{r}
#| label: session-info
#| include: false

sessionInfo()
```






