---
title: "Analysis Inhaler"
author: "Joana"
date: today
toc: true
abstract-title: "Research question:"
abstract: "Does a new inhaled corticosteroid (ICS-X) perform better in a Randomized control trial over 6 weeks in patients with mild persistent asthma compared to a standard inhaler?"
format:
  html:
    embed-resources: true
    number-sections: true
    code-fold: show
editor: visual
editor_options:
  chunk_output_type: console
execute:
  warning: false
  message: false
---

## Prepare session
In this step the Quarto project has been prepared for the analysis of the data set lung function which contains 180 data points of distinct patients. The data set displays information about the age, FEV-1 improvement, baseline FEV-1, treatment group and smoking status. 

```{r}
#| label: prepare-session
#| echo: false

# Load packages
  library(pastecs)
  library(ggpubr); theme_set(theme_bw())
  library(GGally)
  library(performance)
  library(MuMIn)
  library(ggeffects)

  library(knitr) 

setwd("/Users/joanawick/Desktop/BME-321-FS2026-B")

# Import data
lung <- read.csv("data/processed/lung-function.csv")

```


## Explore the data
To look at the different inhalation methods, the dataset is analyzed numerically and graphically.

### Numerically
For the numerical exploration of the data, first the tabel @tbl-explore-data has been created, with which the descriptive statistics can be looked at. 
```{r}
#| label: tbl-explore-data
#| tbl-cap: "Descriptive statistics table of the variables of interest."
#| echo: false

# Numerically
 sum_stats<-
   stat.desc(lung [, c("fev1_improvement", "age", "baseline_fev1", "treatment_group")], norm= T)

 kable(sum_stats, digits= 1)


```

To get information about the proportion in which the various treatments have been used, the tabel @tbl-treatment-group has been created.
```{r}
#| label: tbl-treatment-group
#| tbl-cap: "This table displays information about the proportion of the treatment groups."
#| echo: false

## numerically
table(lung$treatment_group)
```

Another important factor in the analysis is the smoking status. The @tbl-smoking-status serves to further elaborate the smoking status of the participants.
```{r}
#| label: tbl-smoking-status
#| tbl-cap: "This table displays information about the proportion of the participants which smoke."
#| echo: false

## numerically
table(lung$smoking_status)
```


### Graphically 
To get a further impression of the data, it can be displayed graphically. The first step is to compare the treatment groups with @fig-boxplot.

```{r}
#| label: fig-boxplot
#| fig-cap: "This figure shows the outcome between the treatment groups."
#| echo: false 

ggplot(lung, aes(x = treatment_group, y = fev1_improvement)) +
  geom_boxplot() +
  labs(title = "Improvement by Treatment Groups",
    x = "Treatment Group", y = "FEV1 Improvement")

```

With the second @fig-histogram, the distribution of the FEV1 improvement can be explored.
```{r}
#| label: fig-histogram
#| fig-cap: "This figure shows how the FEV1 improvement is distributed."
#| echo: false 

ggplot(lung, aes(x = fev1_improvement)) +
  geom_histogram(bins = 20) +
  labs(title = "Distribution FEV1 Improvement",
    x = "FEV1 Improvement", y = "Count")

```

