---
title: "Analysis Inhaler"
author: "Joana"
date: today
toc: true
abstract-title: "Research question:"
abstract: "Does a new inhaled corticosteroid (ICS-X) perform better in a Randomized control trial over 6 weeks in patients with mild persistent asthma compared to a standard inhaler?"
format:
  html:
    embed-resources: true
    number-sections: true
    code-fold: show
editor: visual
editor_options:
  chunk_output_type: console
execute:
  warning: false
  message: false
---

## Prepare session
In this step the Quarto project has been prepared for the analysis of the data set lung function which contains 180 data points of distinct patients. The data set displays information about the age, FEV-1 improvement, baseline FEV-1, treatment group and smoking status. 

```{r}
#| label: prepare-session
#| echo: false

# Load packages
  library(pastecs)
  library(ggpubr); theme_set(theme_bw())
  library(GGally)
  library(performance)
  library(MuMIn)
  library(ggeffects)

  library(knitr) 

setwd("/Users/joanawick/Desktop/BME-321-FS2026-B")

# Import data
lung <- read.csv("data/processed/lung-function.csv")

```


## Explore the data
To look at the different inhalation methods, the dataset is analyzed numerically and graphically.

### Numerically
```{r}
#| label: tbl-explore-data
#| tbl-cap: "Descriptive statistics table of the variables of interest."
#| echo: false

# Numerically
 sum_stats<-
   stat.desc(lung [, c("fev1_improvement", "age", "baseline_fev1", "treatment_group")], norm= T)

 kable(sum_stats, digits= 1)



```


### Graphically 


